\section{Lecture 17: 04/05/21}

Recall:
\begin{itemize}
    \item A sequence $(a_n)$ in $\R$ converges to a limit $L$ if, for every $\varepsilon > 0$, there exists $N$ such that, for all $n > N$, $|a_n - L| < \epsilon$.
\end{itemize}

\subsection{Monotonic Sequence Property}

\begin{theorem}[Monotonic Sequence Property]
If $(a_n)$ is an increasing sequence which is bounded above, then $(a_n)$ converges (to a limit in $\R$).
\end{theorem}

We have an immediate corollary of this property:
\begin{corollary}
If $(a_n)$ is a decreasing sequence which is bounded below, then $(a_n)$ converges.
\end{corollary}

\begin{proof}
Let $(a_n)$ be a decreasing sequence bounded below by $M$. Then $(-a_n)$ is an increasing sequence bounded above by $-M$, so $(-a_n)$ converges by the Monotonic Sequence Property. Then $(a_n)$ converges by the Algebraic Limit Theorem.
\end{proof}

\subsection{Boundedness}

\begin{definition}[Boundedness]
A set or sequence in $\R$ is \textit{bounded} if there exists $M$ such that $|x| \le M$ for all $x$ in the set or sequence. (This is equivalent to being both bounded above and bounded below).
\end{definition}

\begin{example}
Prove every convergent sequence is bounded.
\end{example}
\begin{proof}
Let $(a_n)_{n=1}^\infty$ be a convergent sequence and $L = \lim_{n=1}^\infty a_n$. Since $1>0$, there exists $N\in \N$ such that, for all $n>N$, $|a_n-L|<1$. Let $M=max(|a_1|,|a_2|,\dots,|a_N|, 1+|L|)$. Let $n\in N$. 
\begin{itemize}
    \item If $n\leq N$, then $|a_n| \leq M$ by definition of $M$.
    \item If $n>M$, then $|a_n-L|<1$, so
    \begin{align*}
        |a_n| &= |a_n-L+L|\\
        &\leq |a_n-L|+|L| \quad\quad \text{by the triangle inequality}\\
        &< 1+|L|\\
        &\leq M
    \end{align*}
\end{itemize}
So, $|a_n| \leq M$.
\end{proof}

\begin{definition}[Subsequence]
Let $(a_n)$ be a sequence. If $n_1 < n_2 < n_3 < \dots$, then $(a_{n_1}, a_{n_2}, a_{n_3}, \dots)$ is a \textit{subsequence} of $(a_n)$.
\end{definition}

\begin{proposition}
If a sequence $(a_n)$ converges to some limit $L$, then every subsequence also converges to $L$.
\end{proposition}

\begin{theorem}[Bolzano-Weierstrass]
Every bounded sequence has a convergent subsequence.
\end{theorem}